---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data

```{r loading_preprocessing of data, echo=TRUE}

## Load ggplot2 library
library(ggplot2)
## Load scales library
library(scales)

## Load the data (i.e. read.csv())
data <- read.csv("activity.csv")

# Process/transform the data (if necessary) into a format suitable for your analysis
total <- aggregate(data$steps, list(date=data$date), sum)
total$steps <- total$x
total$date <- as.Date(total$date)
total$days <- paste("d", seq(1,61), sep="")
avg <- aggregate(x=data$steps, list(date=data$date), mean)

```

## What is mean total number of steps taken per day?
```{r}


# For this part of the assignment, you can ignore the missing values in the dataset.
avg.steps <- mean(data$steps, na.rm=TRUE)

# Make a histogram of the total number of steps taken each day
datebreaks <- as.Date(total$date)
p <- ggplot(total, aes(y=steps, x=date)) +geom_bar(stat="identity")
p <- p + theme(axis.text.x = element_text(angle=90, hjust=1))
p <- p + scale_x_date(breaks=datebreaks, labels=date_format("%b %d") )
# Calculate and report the mean and median total number of steps taken per day

p
```


## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
## Create interval bins of 1-288 (intervals per day)
data$intervalbin <- rep(seq(1,288))

## Calculate the average number of steps per interval
avginterval <- aggregate(x=data$steps, list(y=data$intervalbin), mean, na.rm=TRUE)

## Interval with the maximum number of steps
max.step <- avginterval$y[which.max(avginterval$x)]

## Time series plot of interval vs avg number of steps
plot(avginterval$y, avginterval$x, type="l", xlab="Interval #", ylab="Avg. Steps")
```

Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
## Number of rows with missing values
num.rows.missing <- nrow(data[is.na(data$steps)==TRUE,])

## Missing days/intervals
daysNA <- nrow(array(unique(data$date[is.na(data$steps)])))

## Create a new data set
## Copy the data frame
newdata <- data

## Assign mean steps for each missing interval using means calculated above
newdata[is.na(newdata$steps)==TRUE,1] <- avginterval$x

```


Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}

total2 <- aggregate(newdata$steps, list(date=newdata$date), sum)
total2$steps <- total2$x
total2$date <- as.Date(total2$date)

datebreaks <- as.Date(total2$date)
p2 <- ggplot(total2, aes(y=steps, x=date)) +geom_bar(stat="identity")
p2 <- p2 + theme(axis.text.x = element_text(angle=90, hjust=1))
p2 <- p2 + scale_x_date(breaks=datebreaks, labels=date_format("%b %d") )

```

Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}

```


----------------------------------------------------------------
## Loading and preprocessing the dat
```{r setoptions,echo=TRUE}
library(ggplot2)
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
qplot(data$steps)

stepsmean <- mean(data$steps, na.rm=TRUE)
stepsmedian <- median(data$steps, na.rm=TRUE)
```

The mean number of steps taken per day is: `r stepsmean`.

The median number of steps taken per day is: `r stepsmedian`.


## What is the average daily activity pattern?

```{r}

avginterval <- aggregate(steps~interval, data=data, mean, na.rm=TRUE)
maxinterval <- avginterval$interval[avginterval$steps==max(avginterval$steps)]
qplot(interval, steps, data=newdata, geom="line")
```

## Imputing missing values
```{r}
steps.na <- nrow(data[is.na(data$steps),])
## Create new dataset
newdata <- data
## Create index list of NA steps
na.steps.list <- which(is.na(newdata$steps)==TRUE)
## Replace NA elements with average steps from previous avginterval values
newdata$steps[na.steps.list] <- replace(newdata$steps, na.steps.list, avginterval$steps[avginterval$interval==newdata$interval])
## Create histogram
qplot(newdata$steps)
## Measure new mean
newstepsmean <- mean(newdata$steps)
## Measure new median
newstepsmedian <- median(newdata$steps)
```

The mean number of steps taken per day Without missing values is: `r newstepsmean`.

The median number of steps taken per day without missing values is: `r newstepsmedian`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## Setup a new factor and classify by Weekday and Weekend
newdata$day <- factor(weekdays(as.Date(newdata$date)))
levels(newdata$day)[1:2] <- "Weekday"
levels(newdata$day)[4:6] <- "Weekday"
levels(newdata$day)[2:3] <- "Weekend"
qplot(interval, steps, data=newdata, geom="line", facets=day~.)
```


